---
title: "Decades-Long Decline in Entrepreneurial Dynamism (Part 2)"
subtitle: "A Look at Differences Across Industries"
author: "Asif Mehedi"
date: 2019-04-02
slug: "decline-entrepreneurship-cities"
tags: [entrepreneurship, industries, longitudinal]
---

```{r setup, include=FALSE}
library(tufte)
library(DT)

source(here::here('helpers/setup.R'))
source(here::here('helpers/apiACS.R'))
source(here::here("helpers/drawMaps.R"))

knitr::opts_chunk$set(
  # warning = FALSE,
  # error = FALSE,
  # message = FALSE,
  echo = FALSE,
  fig.width = 8,
  fig.height = 6
)
```

# General trends


```{r}
df_f_all <- read_csv(file = here::here('data_/readonly/bds_f_all_release.csv'))
```

```{r}
get_all_trend <- function(param) {
  ggplot(df_f_all, aes(x = year2, y = param)) + 
    geom_line() + geom_point()
}
```

```{r}
df_f_all %>% 
  select(year2, firms, estabs) %>% 
  gather(firms_or_estabs, number, ... = c(firms, estabs)) %>% 
  ggplot(aes(x = year2, y = number, fill = firms_or_estabs)) + 
    geom_col() + geom_point() + 
    scale_y_continuous(limits = c(0, 7e6), labels = comma)
```

```{r}
# get_all_trend(df_f_all$net_job_creation_rate)

ggplot(df_f_all, aes(x = year2, 
                     y = net_job_creation_rate, 
                     fill = net_job_creation_rate > 0)) + 
  geom_col()
```


```{r}
df_f_all %>% 
  gather(key = creation_destruction, value = number_of_jobs, ... = c(job_creation, job_destruction, net_job_creation)) %>% 
  ggplot(aes(x = year2, y = number_of_jobs, color = creation_destruction)) + 
    geom_line() + geom_point() +
    scale_y_continuous(labels = unit_format(scale = 1/1e6)) +
    geom_abline(slope = 0, intercept = 0, color = "#00ff00")
```

todo: normalize by population size


# Industry trends

## Loading data

todo: add data
  - population for each msa for entire period, either actual or imputed. Get county populations, then roll them up to msa level.

<!-- todo: select the top 3 or 4 sic industries -->

```{r}
df_f_sic <- read_csv(file = here::here("data_/readonly/bds_f_sic_release.csv"))
```

```{r}
# param is the variable we want to see the trend of
get_sic_trend <- function(param = firms, sic_filter = c("07", "10", "15", "20", "40", "50", "52", "60", "70")) {
  sic_filter <- enquo(sic_filter)
  param <- enquo(param)
  df_f_sic %>% 
    select(year2, sic1, !!param) %>% 
    filter(sic1 %in% !!sic_filter ) %>% 
    ggplot(mapping = aes(x = year2, y = !!param, color = sic1)) + 
      geom_line(size = 1) + geom_point(size = 2) +
      scale_y_continuous(labels = scales::comma, limits = c(0, NA))
}
```

## Firms by industry

```{r}
get_sic_trend()

df_f_sic
```


```{r}
get_sic_trend(param = firms, sic_filter = c("70", "52", "15", "60"))
```


## Employment by industry

```{r}
get_sic_trend(param = emp, sic_filter = c("70", "52", "15", "60"))
```


## Job destruction by industry

```{r}
get_sic_trend(param = job_destruction, sic_filter = c("70", "52", "15", "60"))
```

## Job creation by industry

```{r}
get_sic_trend(param = job_creation, sic_filter = c("70", "52", "15", "60"))
```


## Net job creation by industry

### Service industry

```{r}
df_f_sic %>% 
  filter(sic1 %in% c("70")) %>% 
  ggplot(aes(x = year2, y = job_creation, fill = net_job_creation > 0)) +
    geom_col() +
    scale_y_continuous(labels = comma)
```


```{r}
df_f_sic %>% 
  filter(sic1 %in% c("70")) %>% 
  ggplot(aes(x = year2, y = net_job_creation, fill = net_job_creation > 0)) +
    geom_col() +
    scale_y_continuous(labels = comma)
```

The picture becomes most clear when we look at the rate of net job creation, defined as "...".

```{r}
df_f_sic %>% 
  filter(sic1 %in% c("70")) %>% 
  ggplot(aes(x = year2, y = net_job_creation_rate, fill = net_job_creation_rate > 0)) +
    geom_col()
```



```{r}
df_f_sic %>% 
  filter(sic1 %in% c("70")) %>% 
  ggplot(aes(x = year2, y = job_destruction, fill = net_job_creation > 0)) +
    geom_col() +
    scale_y_continuous(labels = comma)
   
```


## Firm exits by industry

```{r}
get_sic_trend(firmdeath_firms)
```

